---
title: "Farah_plots_initial"
output: html_document
date: "2024-07-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
# Read in data
DoctorContacts <- read.csv("UpdatedDoctorContacts.csv")
head(DoctorContacts)
```

```{r}
# Load packages
library(tidyverse)
library(ggplot2)
```
Get column names:

```{r}
names(DoctorContacts)
```

(Visualization): 
Does an individual’s number of chronic diseases (numerical) affect their self-rate health score (categorical)? 
Null: An individual’s number of chronic diseases (num_disease) has no effect on their self-rate health score (health).
Alt: An individual’s number of chronic diseases (num_disease) has an effect on their self-rate health score (health).

```{r}
DoctorContacts$health <- as.factor(DoctorContacts$health)
head(DoctorContacts)
```


```{r}
ggplot(DoctorContacts, aes(x=health, y=num_disease, fill=health)) + geom_boxplot() 
```

(Multilinear Model): 
Does the household head’s years of schooling (schooling), family income (income), family size (family),  statistically significant predictors of deductible rate (deductible).
Null: None of the predictors among the household head’s years of schooling (schooling), family income (income), and family size (family) are helpful in predicting deductible rate (deductible).
Alt: At least one of the predictors among the household head’s years of schooling (schooling), family income (income), and family size (family) are helpful in predicting deductible rate (deductible).

create a model:

```{r}
deductibleAffected_model <- lm(deductible ~ schooling + income + family, data = DoctorContacts)
coef(deductibleAffected_model)
```

```{r}
summary(deductibleAffected_model)
```

```{r}
0.05/3
```

compare the p values to corrected alpha^ 

schooling   0.00116 < alpha
income      < 2e-16 < alpha
family      0.09696 > alpha

therefore family looks sus and maybe we need to drop it 

```{r}
anova(deductibleAffected_model)
```
check linearity assumptions:

```{r}
exp_pred <- mutate(DoctorContacts, predictions=fitted(deductibleAffected_model), residual=residuals(deductibleAffected_model))
head(exp_pred)
```

```{r}
exp_pred <- log_pred[-c(1:11, 13:18)]
tail(exp_pred)
```

```{r}
ggplot(exp_pred, aes(x=predictions, y=residual))+geom_point()+
  geom_hline(yintercept=0, color='red')
```
```{r}
ggplot(exp_pred, aes(sample=residual))+stat_qq() +stat_qq_line()
```
```{r}
df_for_cor <- exp_pred[,c(-5, -6)]
cor_mat<-round(cor(df_for_cor), 2)
cor_mat
```

```{r}
library(ggcorrplot)
ggcorrplot(cor_mat, lab=T, type='lower')
```
there is no fear of clinearity. 

check the AIC:

```{r}
library(MASS)
aic<-stepAIC(deductibleAffected_model, direction='both')
```
```{r}
library(car)
vif(aic)
```



although the vif values look fine, its a little sus because the p value for family is larger than the adjusted value. verify this with prof/others


```{r}
avPlots(deductibleAffected_model)
```

Income  and schooling seem to have an effect. Family is not discernible by the graph. The VIFs show there is not co-linearity. 


The model: 
max_deductible = 419.52262482 * 10.16056417(schooling) + 0.03535185(family) + 7.55246798(income)


visualizing all variables (in curiosity):

 [1] "X"                  "visits"             "log_coinsurance"    "idp"                "log_api"           
 [6] "log_max_deductible" "physlim"            "num_disease"        "health"             "log_income"        
[11] "log_fam_size"       "schooling"          "age"                "sex"                "child"             
[16] "black"              "insurance"          "payment"            "deductible"         "income"            
[21] "family"  

```{r}
long<-gather(DoctorContacts, key='predictor', value='value', visits,
log_coinsurance, idp, physlim, health, num_disease, log_income, log_fam_size, schooling,  
age, sex, child, black, insurance, payment, income, family, log_max_deductible, deductible)
ggplot(long, aes(x=value, y=log_api, color=predictor))+geom_point()+
  facet_wrap(~predictor, scales='free_x')
```


