---
title: "Farah MLR"
output: html_document
date: "2024-08-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


```{r}
library(readr)
library(tidyverse)
library(ggplot2)
library(ggcorrplot)
library(car)
# library(MASS)   # messes with select function in dplyr
library(broom)
library(dplyr)
library(caret)
library(glmnet)
library(pls)
```

```{r}
# Read in data
DoctorContacts <- read.csv("UpdatedDoctorContacts.csv")
head(DoctorContacts)
```

************************************************THE QUESTION***********************************************
(Multilinear Model): 
Does the household head’s years of schooling (schooling), family income (income), family size (family),  statistically significant predictors of deductible rate (deductible).
Null: None of the predictors among the household head’s years of schooling (schooling), family income (income), and family size (family) are helpful in predicting deductible rate (deductible).
Alt: At least one of the predictors among the household head’s years of schooling (schooling), family income (income), and family size (family) are helpful in predicting deductible rate (deductible).

create a model:

predictors: 
  schooling (NUM)
  log_fam_size (NUM)
  log_coinsurance (NUM)
  log_income (NUM)
  physlim (CAT)
  
response:
  deductible (NUM)

```{r}
modeldeductible <- lm(deductible~schooling+log_fam_size+log_coinsurance+log_income+physlim, data=DoctorContacts)
summary(modeldeductible)
```



```{r}
aic <- MASS::stepAIC(modeldeductible, direction = "both", Trace = F)
summary(aic)
```

```{r}
coef(modeldeductible)
```


```{r}
preds <- gather(DoctorContacts, key = "predictor", value = "value", schooling,log_fam_size,log_coinsurance,log_income,physlim)

ggplot(preds, aes(x = value, y = deductible, color = predictor)) + geom_point() +
  facet_wrap(~predictor, scales = "free_x")
```

```{r}
predslasso <- gather(DoctorContacts, key = "predictor", value = "value", schooling,log_fam_size,log_coinsurance,log_income,physlim)

ggplot(predslasso, aes(x = value, y = deductible, color = predictor)) + geom_point() +
  facet_wrap(~predictor, scales = "free_x")
```


```{r}
exp_pred1 <- mutate(DoctorContacts, predictions=fitted(model0), residual=residuals(model0))
exp_pred1 <- log_pred[-c(1:2, 4:6, 8:9, 12:18, 20:21)]
head(exp_pred1)
```


```{r}
ggplot(exp_pred1, aes(x=predictions, y=residual))+geom_point()+
  geom_hline(yintercept=0, color='red')
```

```{r}
ggplot(exp_pred1, aes(sample=residual))+stat_qq() +stat_qq_line()
```
```{r}
df_for_cor <- exp_pred1[,c(-6, -7)]
cor_mat<-round(cor(df_for_cor), 2)
cor_mat
```
```{r}
library(ggcorrplot)
ggcorrplot(cor_mat, lab=T, type='lower')
```
```{r}
vif(modeldeductible)

```


```{r}
avPlots(modeldeductible)
```
```{r}
colnames(DoctorContacts)
```

```{r}
cleaned_contacts <- dplyr::select(DoctorContacts, -visits, -idp, -log_api, -log_max_deductible, 
                                  -num_disease, -health, -age, -sex, -child, -black, -payment, 
                                  -income, -family, -insurance, -X)
cleaned_contacts
```

```{r}
unduplicate_data <- cleaned_contacts[!duplicated(cleaned_contacts), ]
unduplicate_data
```


```{r}
modeldeductible1 <- lm(deductible~., data=unduplicate_data)
summary(modeldeductible1)
```
```{r}
#exp_pred1 <- mutate(unduplicate_data, predictions=fitted(model0), residual=residuals(model0))

```

```{r}
#final_multi_coinsurance <- lm(insurance~., multidata_clean_edited)
#pred_data <- multidata_clean_edited[1:2,-4]

# Prediction interval
#predict(final_multi_coinsurance, newdata = pred_data, interval = "prediction", level = 0.95)
```


```{r}
modeleverything <- lm(deductible~.-X, DoctorContacts)
summary(modeleverything)
```


```{r}
aic <- MASS::stepAIC(modeleverything, direction = "both", Trace = F)
summary(aic)
```
```{r}
DoctorContacts <- dplyr::select(DoctorContacts, -log_max_deductible)
```



```{r}
# look at all x's without y's, drop all 1's in the design matrix so can use glmnet
X_mat<-model.matrix(deductible~0+., data=DoctorContacts)
# don't have all the pts, but have all the predictors
Y<-DoctorContacts$deductible

Lmodel_visits<-glmnet(x=X_mat, y=Y, alpha=1)
# not plot:
plot(Lmodel_visits, label=TRUE, xvar="lambda")
```

#
```{r}
L_visits_glmnet<-cv.glmnet(x=X_mat,y=Y, alpha=1, nfolds=10)
L_visits_glmnet$lambda.1se
```
#
```{r}
plot(Lmodel_visits, label=T, xvar="lambda")+abline(v=log(L_visits_glmnet$lambda.1se))
```

```{r}
DoctorContacts
```

log_coinsurance,idp,physlim, income



```{r}
lassomodel <- lm(deductible~log_coinsurance+idp+physlim+income, DoctorContacts)
summary(lassomodel)
```
```{r}
coef(lassomodel)
```

```{r}
DoctorContacts[81:82,]
```

```{r}
exp(3.258096)
```
```{r}
exp(8.934221)
```


```{r}
lassomodel <- train(deductible~log_coinsurance+idp+physlim+income, method='lm', DoctorContacts)
summary(lassomodel)
```

```{r}
lassomodel$results$RMSE
```
```{r}
summary(DoctorContacts$deductible)
```



```{r}
vif(lassomodel)
```

```{r}
colnames(DoctorContacts)
```
log_coinsurance,idp,physlim, income, deductible


```{r}
data_for_lasso <- dplyr::select(DoctorContacts, -X, -visits, -log_api, -num_disease, -health,
                                -log_income,-log_fam_size, -schooling, -age,-sex,-child,
                                -black,-insurance,-payment,-family)
data_for_lasso
```

```{r}
exp_predlasso <- mutate(data_for_lasso, predictions=fitted(lassomodel), residual=residuals(lassomodel))
exp_predlasso
```



```{r}

ggplot(exp_predlasso, aes(x=predictions, y=residual)) + 
  geom_point() +
  geom_hline(yintercept = 0, color="red")
```

#
```{r}
#QQ plot for normality assumption:
ggplot(exp_predlasso, aes(sample=residual)) +
  stat_qq() +
  stat_qq_line()
```

# Check multicolinearity

```{r}
car::vif(lassomodel)
```
# Make prediction
```{r}
#use predict function for that row=2
predict(lassomodel,exp_predlasso[1:2,-c(4,6,7)] )

pred_data <- exp_predlasso[81:82,-c(4,6,7)] 

predict(lassomodel, newdata = pred_data, interval = "prediction", level = 0.95)

```
SO BAD
```{r}
# Confidence interval
predict(lassomodel, newdata = pred_data, interval = "confidence", level = 0.95)
```

SO BAD


```{r}

```

