---
title: "Farah MLR"
output: html_document
date: "2024-08-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


```{r}
library(readr)
library(tidyverse)
library(ggplot2)
library(ggcorrplot)
library(car)
# library(MASS)   # messes with select function in dplyr
library(broom)
library(dplyr)
library(caret)
library(glmnet)
library(pls)
```

```{r}
# Read in data
DoctorContacts <- read.csv("UpdatedDoctorContacts.csv")
head(DoctorContacts)
```

************************************************THE QUESTION***********************************************
(Multilinear Model): 
Does the household head’s years of schooling (schooling), family income (income), family size (family),  statistically significant predictors of deductible rate (deductible).
Null: None of the predictors among the household head’s years of schooling (schooling), family income (income), and family size (family) are helpful in predicting deductible rate (deductible).
Alt: At least one of the predictors among the household head’s years of schooling (schooling), family income (income), and family size (family) are helpful in predicting deductible rate (deductible).

create a model:

predictors: 
  schooling (NUM)
  log_fam_size (NUM)
  log_coinsurance (NUM)
  log_income (NUM)
  physlim (CAT)
  
response:
  deductible (NUM)

```{r}
modeldeductible <- lm(deductible~schooling+log_fam_size+log_coinsurance+log_income+physlim, data=DoctorContacts)
summary(modeldeductible)
```



```{r}
aic <- MASS::stepAIC(modeldeductible, direction = "both", Trace = F)
summary(aic)
```

```{r}
coef(modeldeductible)
```

```{r}
colnames(DoctorContacts)
```


```{r}
preds <- gather(DoctorContacts, key = "predictor", value = "value", schooling,log_fam_size,log_coinsurance,log_income,physlim)

ggplot(preds, aes(x = value, y = deductible, color = predictor)) + geom_point() +
  facet_wrap(~predictor, scales = "free_x")
```

```{r}
exp_pred1 <- mutate(DoctorContacts, predictions=fitted(model0), residual=residuals(model0))
exp_pred1 <- log_pred[-c(1:2, 4:6, 8:9, 12:18, 20:21)]
head(exp_pred1)
```


```{r}
ggplot(exp_pred1, aes(x=predictions, y=residual))+geom_point()+
  geom_hline(yintercept=0, color='red')
```

```{r}
ggplot(exp_pred1, aes(sample=residual))+stat_qq() +stat_qq_line()
```
```{r}
df_for_cor <- exp_pred1[,c(-6, -7)]
cor_mat<-round(cor(df_for_cor), 2)
cor_mat
```
```{r}
library(ggcorrplot)
ggcorrplot(cor_mat, lab=T, type='lower')
```
```{r}
vif(modeldeductible)

```

