---
title: "DevonCodeFP"
author: "DH"
date: "2024-07-25"
output: html_document
---
Data cleaning
```{r}
library(tidyverse)
library(ggplot2)
library(caret)
library(pls)
```

```{r}
#setwd("C:/Users/chris/Documents/UVA/STAT_6021/FP")
data<-read.csv('DoctorContacts.csv')
View(data)
```

```{r}
DoctorContacts <- data %>%
  rename(visits = mdu, log_coinsurance = lc, log_api = lpi,
         log_max_deductible = fmde, num_disease = ndisease,
         log_income = linc, log_fam_size = lfam, schooling= educdec)
```

```{r}
DoctorContacts$insurance = exp(DoctorContacts$log_coinsurance)
DoctorContacts$payment = exp(DoctorContacts$log_api)
DoctorContacts$deductible = exp(DoctorContacts$log_max_deductible)
DoctorContacts$income = exp(DoctorContacts$log_income)
DoctorContacts$family = exp(DoctorContacts$log_fam_size)
View(DoctorContacts)
```
The most updated version of the csv is now called DrContactsNew. It includes
log and sqrt transformations as well as the original raw data(that we had to
untransform from the log).
```{r}

```
Data Visualization

Null hypothesis = There is no relationship between mean health score and education
Alternative hypothesis = At least one of the means is different from the others

```{r}
DoctorContacts$health <- factor(DoctorContacts$health, 
                                levels = c('poor', 'fair', 'good', 'excellent'))
ggplot(DoctorContacts, aes(x=health, y=schooling, fill=health)) + geom_boxplot()+ ggtitle("Years of Education vs. Self-Reported Health Score") + theme(plot.title = element_text(hjust = 0.5))+ theme(legend.position="none")+labs(x = "Self-Rated Health", y = "Schooling (Years)")


```

```{r}
anova <- aov(schooling~health, data=DoctorContacts)
summary(anova)
```
There is an extremely small p-value here, which provides strong evidence to
reject the null hypothesis that the family head's years of schooling have no 
significant difference on how patients self report their own health.
One thing to note is the very high SSE value (146801). This means that most of
the variation in the response cannot be explained by a linear regression model.
This makes sense because since one of these values is categorical, it would be
difficult to represent it as a linear regression model.

```{r}
TukeyHSD(anova, conf.level = 0.95)
```
These results show that all tests are statistically significant. The mean of
schooling of people who reported their own health as excellent is higher than
those who reported good, fair or poor in that order.
This data is observational, so we can not make cause and effect statements,
but we are 95% confident...

```{r}
plot(TukeyHSD(anova))
```
Logistic Regression
```{r}
drs<-read.csv('DrContactsNew.csv')
View(drs)
```

Transform health column into binary as column 'condition'
Used log data from original data set for this
```{r}
 drs1<- DoctorContacts %>%
  mutate(condition = case_when(
    health %in% c("poor", "fair") ~ "unhealthy",
    health %in% c("good", "excellent") ~ "healthy",
    TRUE ~ as.character(health)
  ))

View(drs1)
```

Bar graph shows relationship between condition and sex. (Just for fun)
```{r}
ggplot(drs1, aes(x=sex, fill=factor(condition), color=factor(condition)))+
  geom_bar(position='fill')
```


Make all categoricals factors (0,1)
```{r}
drs2<-drs1[, -c(1,9)]%>%
  mutate(physlim=ifelse(physlim=="TRUE", 1,0),
         idp=ifelse(idp=="TRUE", 1,0),
         sex=ifelse(sex=="male", 1,0),
         child=ifelse(child=="TRUE", 1,0),
         black=ifelse(black=="TRUE", 1,0),
         condition=ifelse(condition=="healthy", 1,0))
View(drs2)
```

Remove response/unneeded variables, make correlation matrix
```{r}
dat<-drs2[, -15]
cor_mat<-round(cor(dat),2)
ggcorrplot::ggcorrplot(cor_mat, lab=T, type='lower')
```
There is multicollinearity between age and child (to be expected) and
log_max_deductible and log_coinsurance (also to be expected). I will remove
log_max_deductible and child.

```{r}
dat2<-dat[, -c(5,13)]
cor_mat2<-round(cor(dat2),2)
ggcorrplot::ggcorrplot(cor_mat2, lab=T, type='lower')
```

```{r}
drs3<-drs2[, -c(5,13)]
```

Create logistic regression model using glm
```{r}
logit_model1<-glm(condition~., drs3, family='binomial')
summary(logit_model1)
```
It looks like family size, coinsurance, idp, and sex are not signficant,
so I will remove them in the next iteration.

```{r}
logit_model2<-glm(condition~.-idp-log_coinsurance-log_fam_size-sex, drs3, family='binomial')
summary(logit_model2)
```
Yay! Check with AIC
```{r}
aic<-MASS::stepAIC(logit_model1, direction='both', trace=F)
summary(aic)
```
Matches logit_model2, yay!

```{r}
car::vif(logit_model2)
```
Low VIF suggests no multicollinearity (also shown by correlation matrix?) Yay!

```{r}
#visualization of proportion of black folks in dataset

black_count <- sum(DoctorContacts$black)
non_black_count <- sum(!DoctorContacts$black)
total_count <- nrow(DoctorContacts)

black_count
non_black_count
total_count
```
```{r}
counts <- c(black_count, non_black_count)
totals <- c(total_count, total_count)

prop.test(counts, totals)
```
```{r}
drs4 <- DoctorContacts %>%
  mutate(black = ifelse(black == "TRUE", 
                        "Black Head of Household", 
                        ifelse(black == "FALSE", 
                               "Non-Black Head of Household", 
                               black)))

ggplot(drs4, aes(x=black, y=visits, fill=black))+geom_boxplot()+
labs(x = "", y = "Number of Annual Doctor Visits")+ ggtitle("Number of Doctor Visits vs. Race") + theme(plot.title = element_text(hjust = 0.5))+ theme(legend.position="none")
```
```{r}
means <- drs4 %>%
  group_by(black) %>%
  summarise(mean_visits = mean(visits, na.rm = TRUE))
means
```
```{r}
t.test(visits~black, data=DoctorContacts)

```


```{r}
data<-read.csv('UpdatedDoctorContacts.csv')
View(data)
```

```{r}
coinsurance<-ggplot(data, aes(x=insurance)) + geom_bar(fill='blue')+
labs(x = "Co-Insurance Rate", y = "Count")+ ggtitle("Co-Insurance Rates") + theme(plot.title = element_text(hjust = 0.5))+ theme(legend.position="none")
coinsurance
```
```{r}
data$insurance %>%
  summary(
    median = median(values),
    Q1 = quantile(values, probs = 0.25),
    Q3 = quantile(values, probs = 0.75)
  )
```
```{r}
num_disease<-ggplot(data, aes(y=num_disease)) + geom_boxplot(fill='blue', width=0.5)+
labs(x = "", y = "Number of Chronic Diseases")+ ggtitle("Chronic Diseases") + theme(plot.title = element_text(hjust = 0.5))+ theme(legend.position="none") + xlim(-1,1)
num_disease
```
```{r}
data$num_disease %>%
  summary(
    median = median(values),
    Q1 = quantile(values, probs = 0.25),
    Q3 = quantile(values, probs = 0.75)
  )
```

```{r}
visits<-ggplot(data, aes(x=visits)) + geom_bar(fill='blue')+
labs(x = "Doctor Visits", y = "Count")+ ggtitle("Number of Annual Doctor Visits") + theme(plot.title = element_text(hjust = 0.5))+ theme(legend.position="none")
visits
```
```{r}
data$visits %>%
  summary(
    median = median(values),
    Q1 = quantile(values, probs = 0.25),
    Q3 = quantile(values, probs = 0.75)
  )
```

```{r}
deductible<-ggplot(data, aes(y=deductible)) + geom_boxplot(fill='sky blue')+
labs(x = "", y = "Deductible")+ ggtitle("Deductible Rate") + theme(plot.title = element_text(hjust = 0.5))+ theme(legend.position="none")+ xlim(-1,1)
deductible
```
```{r}
data$deductible %>%
  summary(
    median = median(values),
    Q1 = quantile(values, probs = 0.25),
    Q3 = quantile(values, probs = 0.75)
  )
```

```{r}
income<-ggplot(data, aes(y=income)) + geom_boxplot(fill='sky blue')+
labs(x = "", y = "Annual Income")+ ggtitle("Annual Income") + theme(plot.title = element_text(hjust = 0.5))+ theme(legend.position="none")+xlim(-1,1)
income
```
```{r}
data$income %>%
  summary(
    median = median(values),
    Q1 = quantile(values, probs = 0.25),
    Q3 = quantile(values, probs = 0.75)
  )
```

```{r}
payment<-ggplot(data, aes(y=payment)) + geom_boxplot(fill='sky blue')+
labs(x = "", y = "Payment (USD)")+ ggtitle("Annual Participation Incentive Payment") + theme(plot.title = element_text(hjust = 0.5))+ theme(legend.position="none")+xlim(-1,1)
payment
```

```{r}
data$payment %>%
  summary(
    median = median(values),
    Q1 = quantile(values, probs = 0.25),
    Q3 = quantile(values, probs = 0.75)
  )
```

```{r}

```

