---
title: "Multilinear Model for Coinsurance Prediction"
author: "Kayla Lee"
date: "2024-08-05"
output: html_document
---

Set up
```{r}
library(readr)
library(tidyverse)
library(ggplot2)
library(ggcorrplot)
library(car)
# library(MASS)   # messes with select function in dplyr
library(broom)
library(dplyr)
library(caret)
library(glmnet)
library(pls)

DoctorContacts <- read_csv("DoctorContacts.csv")

# Rename columns
DoctorContacts <- DoctorContacts %>%
  rename(visits = mdu, log_coinsurance = lc, log_api = lpi,
         log_max_deductible = fmde, num_disease = ndisease,
         log_income = linc, log_fam_size = lfam, schooling = educdec)

# Delete extra index column
DoctorContacts <- select(DoctorContacts, -rownames)

# Undo log transformations
DoctorContacts$coinsurance = exp(DoctorContacts$log_coinsurance)
DoctorContacts$payment = exp(DoctorContacts$log_api)
DoctorContacts$deductible = exp(DoctorContacts$log_max_deductible)
DoctorContacts$income = exp(DoctorContacts$log_income)
DoctorContacts$family = exp(DoctorContacts$log_fam_size)
```


AIC MODEL FOR COINSURANCE
```{r}
multi_coinsurance_model <- lm(coinsurance~visits+idp+physlim+num_disease+black+payment+deductible+income+family-log_coinsurance-log_api-log_max_deductible-log_income-log_fam_size, DoctorContacts)
aic <- MASS::stepAIC(multi_coinsurance_model, direction = "both", Trace = F)
summary(aic)
```

```{r}
final_multi_model <- lm(coinsurance~.-log_coinsurance-log_api-log_max_deductible-log_income-log_fam_size, DoctorContacts)
aic <- MASS::stepAIC(final_multi_model, direction = "both", Trace = F)
summary(aic)
```

Despite taking out some additional predictors in the first model, the second multilinear model has slightly higher R^2 and adjusted R^2 values.

```{r}
coef(final_multi_model)
```

```{r}
final_multi_preds <- gather(DoctorContacts, key = "predictor", value = "value", visits, idp, num_disease, 
                            health, schooling, age, sex, child, black, payment, deductible, income, family)

ggplot(final_multi_preds, aes(x = value, y = coinsurance, color = predictor)) + 
  geom_point() +
  facet_wrap(~predictor, scales = "free_x")
```

Since linearity fails we will not move on with this model.







ATTEMPT 2 AT MULTILINEAR MODEL PREDICTING COINSURANCE
------------------------------- MODELS -----------------------------------------
```{r}
# Data
multidata <- read.csv("DrContactsNew.csv")
multidata_clean <- select(multidata, -c(X.1, X, log_coinsurance, sqrt_insurance))

# First model: comparing coinsurance to everything
multi_coinsurance_model1 <- lm(insurance~., multidata_clean)
summary(multi_coinsurance_model1)
```

```{r}
# Second model: hand-selecting the p-values that are significant
multi_coinsurance_model2 <- lm(insurance~.-idp-log_api-health-log_fam_size-schooling-sex-child-black-family-log_max_deductible-log_income-sqrt_payment-sqrt_deductible-sqrt_income-sqrt_family, multidata_clean)
summary(multi_coinsurance_model2)
```

```{r}
# Third model: hand-selecting the p-values that are significant (another attempt)
multi_coinsurance_model3 <- lm(insurance~.-physlim-sex-log_api-health-log_fam_size-schooling-age-child-black-log_max_deductible-log_income-sqrt_payment-sqrt_deductible-sqrt_income-sqrt_family, multidata_clean)

summary(multi_coinsurance_model3)
```



------------------------ CHECKING MODEL ASSUMPTIONS ----------------------------
WE ARE GOING WITH multi_coinsurance_model3
```{r}
# Checking linearity
multi_coinsurance_preds <- gather(multidata_clean, key = "predictor", value = "value", 
                                  visits, idp, num_disease, payment, deductible, income, family)

ggplot(multi_coinsurance_preds, aes(x = value, y = insurance, color = predictor)) + 
  geom_point() +
  facet_wrap(~predictor, scales = "free_x")
```
There does not appear to be any linearity.

```{r}
# Creating a residuals column
multi_coinsurance_model3 <- lm(insurance~.-physlim-sex-log_api-health-log_fam_size-schooling-age-child-black-log_max_deductible-log_income-sqrt_payment-sqrt_deductible-sqrt_income-sqrt_family, multidata_clean)

coinsurance_multi_predictions <- mutate(multidata_clean, predictions = fitted(multi_coinsurance_model3), 
                                        residuals = residuals(multi_coinsurance_model3))


# Plotting residuals
ggplot(coinsurance_multi_predictions, aes(x = predictions, y = residuals)) + 
  geom_point() + 
  geom_hline(yintercept = 0, color = 'red')
```
The residuals are not randomly distributed.

```{r}
# Checking residual normality
ggplot(coinsurance_multi_predictions, aes(sample = residuals)) + 
  stat_qq() + 
  stat_qq_line(color = 'red')
```
Points fall very close to the line, looks pretty good.



----------------------- CHECKING FOR MULTICOLLINEARITY  ------------------------
```{r}
# Creating a correlation matrix to check for multicollinearity
multidata_clean_edited <- select(multidata_clean, -c(log_api, log_max_deductible, physlim, health, log_income,
                                                     log_fam_size, schooling, age, sex, child, black, 
                                                     sqrt_payment, sqrt_deductible, sqrt_income, sqrt_family))


cor_matrix <- round(cor(multidata_clean_edited), 2)
cor_matrix
```

```{r}
# Plotting the correlation matrix to check for multicollinearity
ggcorrplot(cor_matrix, lab = TRUE, type = "lower")
```

```{r}
# Looking at VIF to check for multicollinearity
vif(multi_coinsurance_model3)
```



--------------- LOOKING AT INDIVIDUAL PREDICTORS VS COINSURANCE  ---------------
```{r}
# Individual predictors' impacts on response (coinsurance)
avPlots(multi_coinsurance_model3)
```



---------------------------- MAKING PREDICTIONS --------------------------------
```{r}
# Data for predictions
final_multi_coinsurance <- lm(insurance~., multidata_clean_edited)
pred_data <- multidata_clean_edited[1:2,-4]

# Prediction interval
predict(final_multi_coinsurance, newdata = pred_data, interval = "prediction", level = 0.95)
```

```{r}
# Confidence interval
predict(final_multi_coinsurance, newdata = pred_data, interval = "confidence", level = 0.95)
```









